phoeton: Software for Objective Probabilistic Foehn Classification

### 2018-11-30 inst/draft/draft.tex
Updated draft and refs

Introduction why this is so great. Georg might like to contribute
here. Just to leave/include some references:

\begin{itemize}
    \item finite mixture model \citep{leisch2004,gruen2008}
    \item automatic probabilistic foehn classification \cite{plavcan2014}
    \item IWLS/IRLS \citep{mccullagh1989}
\end{itemize}

% -------------------------------------------------------------------
% SECTION: METHODS
% -------------------------------------------------------------------
\section{Methods}

@@

@@

% -------------------------------------------------------------------
% SECTION: SOFTWARE
% -------------------------------------------------------------------
\section{Software}

>>>>

Introduction why this is so great. Georg might like to contribute
here. Just to leave/include some references:
\begin{itemize}
    \item finite mixture model \citep{leisch2004,gruen2008}
    \item automatic probabilistic foehn classification \cite{plavcan2014}
    \item IWLS/IRLS \citep{mccullagh1989}
    \item Community foehn classification experiment \cite{mayr2018unpublished}
\end{itemize}

I would go for something like:

Gap winds are typically characterized by strong ncrease of wind speed with a rapid
increase in temperature and a drop in relative humidity. Such winds can be
found in various regions on our globe and bear different names for specific
regions such as the Santa-Anna winds in California which became infamous due to
their destructive effect in combination with wilde fires, the Chinook winds in
the Rocky Mountain area, Mistral and Bora in south France and Croatia, or the
Foehn winds in the European Alps.  Over the past decades \textit{foehn wind}
has become a synonym for such gap flows and will be used in this article,
however, the method as presented is applicable for all gap flow winds
independent from their geographical location.

The conceptual model of foehn winds is quite well known
\citep{armi2007,mayr2008,mayr2011}, however, foehn cannot measured directly.
Historically foehn was often diagnosed manually by looking at available
meteorological records such as wind speed and direction, time series of
temperature and humidity, the conditions of the sky (e.g., occurence of
lenticularis clouds in the region), or observations upstream and downstream of
the location of interest.  Some of the longest records of foehn observations
can be found in Switzerland (Altdorf \note{i'm just guessing!}) reaching back
to 1850 (\note{just guessing!}) for a hand full of stations.  These records are
typically based manual classification made by experienced weather observers,
weather enthusiasts, and meteorologists.  This task is time consuming and is
only done for some selected sites, typically sites with long historical
records. To be able to extend the diagnosis and to perform foehn classification
on a large number of meteorological sites computer amed classification is
needed.

Few attepmts to automatically classify foehn winds have been proposed over the
past years such as \note{cite this and that}. These methods are only binary
(either \textit{yes} or \textit{no}) and reqire manual intervention to tune the
hyper-parameters, e.g., thresholds used to classify a specific observation as
foehn. In \citeyear{plavcan2014} \citeauthor{plavcan2014} proposed a new
objective and probabilistic foehn classification method based on a finite
mixture model, a statistical regression-based model for automatic foehn
classification. In a recent study of \cite{mayr2018unpublished} this new method
has been tested against existing methods, students, and experts to to proof the
diagnostic power of the new approach.

Overall, the automatic finite mixture model based foehn classification
\citep{plavcan2014} is able to compete with the human classification and 
has the advantages that it is (i) reproductable, and (ii) can easily be
applied to large volumes of data. This allows one to perform foehn classification
for a wide range of different sites.

The article by \citep{plavcan2014} introduced the method using a powerful
package called \code{FlexMix} \citep{leisch2004,gruen2007,gruen2008} available
for the statistical software \textit{R}.  However, only little guidance was
given by \citep{plavcan2014} on how to apply this method to own data and an
implementation of the method outside the \textit{R} programming language has
was not available until today.
Thus, the aim of this article is to introduce the algorithm behind the
automated probabilistic foehn classification \citep{plavcan2014} in more
detail and introducing new open-source software for \textit{R} and \textit{Python}
which provide the required methods to perform the classification and a set
of diagnostic tools for this specific task.

Section~\ref{sec:methods} describes the method itself followed by
Section~\ref{sec:software}  introducing the software with a set of
examples. Section~\dots \note{case study? Required for this article?}

% -------------------------------------------------------------------
% SECTION: METHODS
% -------------------------------------------------------------------
\section{Methods}\label{sec:methods}

@@

Thus, this section introduces a simplified approach to estimate 
a finite mixture model for the special case with two distinct
Gaussian distributions and a binomial logit model to estimate
the coefficients for the concomitant variables.

\note{\dots ... and so far, and so on.}

@@

% -------------------------------------------------------------------
% SECTION: SOFTWARE
% -------------------------------------------------------------------
\section{Software}\label{sec:software}
